### 2016-03-07 paper/paper.tex

\usetikzlibrary{external} \tikzexternalize[prefix=figures/precompiled/]

    \newlength\figureheight \newlength\figurewidth

@@

\twocolumn[

\aistatstitle{Learning to Control Kilobots with a Flashlight}
\aistatsauthor{Alexander Hendrich \And Daniel Kauth \And Gregor Gebhardt}
\aistatsaddress{TU Darmstadt \And TU Darmstadt \And TU Darmstadt}]

@@

With $n$ being the number of sample and $k$ being the number of features
we define the two feature matrices $\Phi$ and $\Phi'$ (size $n\times k$)
$$
\Phi = \left(
\begin{array}{c}
    \phi(s_1,a_1)^T \\
    \phi(s_2,a_2)^T \\
    \dots \\
    \phi(s_n,a_n^T)
\end{array} \right), \;
\Phi' = \left(
\begin{array}{c}
    \mathbf{E}_{\pi(a_1'|s_1')} [\phi(s_1', a_1') | s_1']^T \\
    \mathbf{E}_{\pi(a_2'|s_2')} [\phi(s_2', a_2') | s_2']^T \\
    \dots \\
    \mathbf{E}_{\pi(a_n'|s_n')} [\phi(s_n', a_n') | s_n']^T \\
\end{array} \right)
$$
and the reward vector
$$
R = \left(
\begin{array}{c}
    r(s_1) \\
    r(s_2) \\
    \dots \\
    r(s_n)
\end{array} \right).
$$
For $\Phi'$ we estimate the expected value
$$
\mathbf{E}_{\pi(a_i'|s_i')} [\phi(s_i', a_i') | s_i']
$$
by taking the average over $p$ samples $a_i' \sim \pi(a_i'|s_i')$, where $\pi$
is our current policy.

Using the derivations of \cite{lspi} and \cite{lstdRegularization} we can easily
compute $\theta$ as
$$
\theta = (X^TX+\beta'I)^{-1}X^Ty
$$
with
$$X = C(A+\beta I) \qquad y=Cb$$
$$A = \Phi^T(\phi-\gamma \Phi') \qquad b=\Phi^T R$$
$$C = \Phi(\Phi^T\Phi+\beta I)^{-1},$$
where $\beta$ and $\beta'$ are regularization parameters for the projection and
fixed-point step, which make the estimation numerically more stable and improve
noise tolerance.

@@

To achieve this we optimize over the joint state action distribution
$p(s,a) = p(s) \pi'(s, a)$ and require that the estimated state distribution
$p(s)$ is the same as in the given state sample distribution $\mu(s)$.
This is implemented by matching feature averages of $p(s)$ and $\mu(s)$ and
leads to the following optimization problem:
\begin{align}
    & \max_p \int p(s, a)Q(s, a)\:ds\:da\:, \\
    & \text{s.t. } \text{KL}(p(s, a) || \mu(s)\pi(a|s)) \leq \epsilon,\\
    & \int p(s)\phi(s)\:ds = \hat{\phi}, \int p(s, a)\:ds\:da = 1,
\end{align}
where $\hat{\phi} = \int p(s) \phi(s)\:ds$ is the average feature vector of all
state samples and $\phi(s)$ is another feature function.

This problem can be solved using Lagrangian multipliers and has the solution
$$
p(s)\pi'(a|s) \propto \mu(s) \pi(a|s) \exp\left(\frac{Q(s,a) - V(s)}{\eta}\right),
$$
where $V(s) = v^T \phi(s)$ and $\eta$ and $v$ are Lagrangian multipliers which
can be obtained by minimizing the dual function using standard gradient methods.

\subsubsection{Obtaining a new Policy}
Lastly we need to fit a new policy to our desired probabilities
$p(s_i, a_i) = p(s_i) \pi'(a_i|s_i)$. This is done by minimizing the KL
divergence between $\pi'$ and our new policy $\bar{\pi}$ which is equivalent to
computing a weighted maximum likelihood estimate with sample weights
$$
w_i = \exp\left(\frac{Q(s_i, a_i) - V(s_i)}{\eta}\right).
$$

In this paper we use a weighted Sparse Gaussian Process to obtain our new policy
$\bar{\pi}$. TODO citation / explain Sparse GP?

@@

\subsection{Object Pushing Policy}

\subsection{Maze solving Policy}

TODO use maze solver or just points list?

\section{FUTURE WORK \& CONCLUSION} Our goal for the second part of this project
is to learn how to control the kilobots, so that they can push a given object
(e.g. flat cylinder) towards the desired direction given by our first policy. To
achieve this goal, the policy first has to move the kilobots behind the object.
Depending on the shape of the object (flat cylinder compared to an ``L''-shaped
object), the optimal positions for the kilobots to push can be different if the
object is rotated.

Our initial approach was to solve the two problems at once, so we already went
through some attempts to model the pushing behavior. Although we could not
achieve satisfactorily results with this approach, we feel confident that our
modeling will return good results once we split up the problem into two distinct
problems. To model the pushing behavior we need to find a good position behind
the object and then push the object with the kilobots. Therefore we placed two
rings of RBFs around the object's position. Because we split up the two policies
we can also increase the number of RBFs we used for the pushing behavior or even
spread them on a grid around the object.

Because we want to apply our learned policies to larger quantities of kilobots,
we will probably have to simulate a large quantity of kilobots during sample
generation. Because the number of simulated kilobots is very crucial to the
performance and time complexity we will have to keep this number quite small.
Maybe we can only simulate a few kilobots or even a single one and let them
represent the center of mass of a whole swarm of kilobots. Otherwise we would
probably have to further improve the simulation performance and extend the
parallel computing capabilities to a larger cluster.

Until now we have learned a policy for moving an object through a maze towards a
goal position. Once the second problem has been solved, we should be able to
easily combine both policies to solve the initial problem of pushing an object
through a maze by controling a swarm of kilobots with a light source. Finally we
want to transfer the resulting policies onto a real world scenario with a self
built maze and real kilobots.

>>>>

\newlength\figureheight \newlength\figurewidth % ?

@@

\twocolumn[
\aistatstitle{Learning to Control Kilobots with a Flashlight}
\aistatsauthor{Alexander Hendrich \And Daniel Kauth \And Gregor Gebhardt}
\aistatsaddress{TU Darmstadt \And TU Darmstadt \And TU Darmstadt}]

@@

With $n$ being the number of sample and $k$ being the number of features
we define the two feature matrices $\Phi$ and $\Phi'$ (size $n\times k$)
$$
\Phi = \left(
\begin{array}{c}
    \phi(s_1,a_1)^T \\
    \phi(s_2,a_2)^T \\
    \dots \\
    \phi(s_n,a_n^T)
\end{array} \right), \;
\Phi' = \left(
\begin{array}{c}
    \mathbf{E}_{\pi(a_1'|s_1')} [\phi(s_1', a_1') | s_1']^T \\
    \mathbf{E}_{\pi(a_2'|s_2')} [\phi(s_2', a_2') | s_2']^T \\
    \dots \\
    \mathbf{E}_{\pi(a_n'|s_n')} [\phi(s_n', a_n') | s_n']^T \\
\end{array} \right)
$$
and the reward vector
$$
R = \left(
\begin{array}{c}
    r(s_1) \\
    r(s_2) \\
    \dots \\
    r(s_n)
\end{array} \right).
$$
For $\Phi'$ we estimate the expected value
$$
\mathbf{E}_{\pi(a_i'|s_i')} [\phi(s_i', a_i') | s_i']
$$
by taking the average over $m$ samples $a_i' \sim \pi(a_i'|s_i')$, where $\pi$
is our current policy.

Using the derivations of \cite{lspi} and \cite{lstdRegularization} we can easily
compute $\theta$ as
$$
\theta = (X^TX+\beta'I)^{-1}X^Ty
$$
with
$$X = C(A+\beta I) \qquad y=Cb$$
$$A = \Phi^T(\phi-\gamma \Phi') \qquad b=\Phi^T R$$
$$C = \Phi(\Phi^T\Phi+\beta I)^{-1},$$
where $\beta$ and $\beta'$ are regularization parameters for the projection and
fixed-point step which make the estimation numerically more stable and improve
noise tolerance.

@@

To achieve this we optimize over the joint state action distribution
$p(s,a) = p(s) \pi'(s, a)$ and require that the estimated state distribution
$p(s)$ is the same as in the given state sample distribution $\mu(s)$.
This is implemented by matching feature averages of $p(s)$ and $\mu(s)$ and
leads to the following optimization problem:
\begin{align*}
    & \max_p \int p(s, a)Q(s, a)\:ds\:da\:, \\
    & \text{s.t. } \text{KL}(p(s, a) || \mu(s)\pi(a|s)) \leq \epsilon,\\
    & \int p(s)\phi(s)\:ds = \hat{\phi}, \int p(s, a)\:ds\:da = 1,
\end{align*}
where $\hat{\phi} = \int p(s) \phi(s)\:ds$ is the average feature vector of all
state samples, $\phi(s)$ is another feature function and $\epsilon$ is the
maximum allowed distance between the old and new policies.

This problem can be solved using Lagrangian multipliers and has the solution
$$
p(s)\pi'(a|s) \propto \mu(s) \pi(a|s) \exp\left(\frac{Q(s,a) - V(s)}{\eta}\right),
$$
where $V(s) = v^T \phi(s)$ and $\eta$ and $v$ are Lagrangian multipliers which
can be obtained by minimizing the dual function on the samples using standard
gradient methods.

\subsubsection{Obtaining a new Policy}
The previous step only gives us desired probabilities
$p(s_i, a_i) = p(s_i) \pi'(a_i|s_i)$ for our samples so in this step we need
to generalize these by fitting a new policy $\bar{\pi}$.

This is done by minimizing the KL divergence between $\pi'$ and our new policy
$\bar{\pi}$ which is equivalent to computing a weighted maximum likelihood
estimate with sample weights
$$ w_i = \exp\left(\frac{Q(s_i, a_i) - V(s_i)}{\eta}\right). $$

In this paper we use a weighted Gaussian Process (GP) to obtain our new policy
$\bar{\pi}$ by estimating action $a_i$ for state $s_i$ with weight $w_i$. We
also use the sparse version of the GP to reduce computation time.

This choice of weights will reproduce actions wich lead to a higher reward with
higher probability but will also choose suboptimal actions with some
probability, because our new policy can not be too far away in terms of KL
divergence from the old policy and we start with a random policy. This achieves
a tradeoff between exploration and exploitation.

TODO citation / explain Sparse GP?

@@

Our learning process is implemented in Python and uses a simple kilobot
simulator (see \autoref{sec:simulator}) written by us to generate samples.

\subsection{Maze Solving Policy}
This policy tells us where the object should move given the current object
position and solves the maze (\autoref{fig:simulator}) in this way.

Because this policy is relatively easy to learn and there are specialized maze
solvers which can easily compute an optimal path through any maze we did not
learn this policy and instead focused on the second one, which is much more
interesting. Therefore we manually specified this policy by providing a number
of waypoints the object should pass and giving the next waypoint as the target
position. Once we reach a waypoint we move one to next one.

In our experiments this simple policy is sufficient to move different objects
through a simple maze using our object movement policy. For this reason we will
only discuss the object moving policy in the following sections.

TODO dont learn first part because already done in first paper?
TODO use maze solver or just points list?

\subsection{Simulator}
\label{sec:simulator}

\newcommand\nkb{\#\mathit{kb}}
\newcommand\nep{\#\mathit{ep}}
\newcommand\nst{\#\mathit{st}}

To simulate the kilobots and collect samples for the object movement policy we
wrote our own simple kilobot simulator (\autoref{fig:simulator}) in Python based
on the 2D physics engine pybox2d\cite{pybox2d}. This allows us to simulate
collisions between kilobots and other kilobots, the object and walls.

When requesting new samples from the simulator we specify the current policy
$\pi$ which will be used to generate these sampels as well as the sampling
parameters such as the shape of the object, the number of kilobots $\nkb$, the
number of episodes $\nep$ and the number of steps per episode $\nst$.

Given these parameters the simulator first generates $\nep$ kilobot start
positions $\mathit{start}_i$ equally spaced in a circle of radius $1.5 \cdot r$
around the object which starts in the middle of the screen where $r$ is the
radius or half width of the object depending on the its shape. Then for
$i \in \{1, \dots, \nep\}$ it does the following:

The object's position is reset to the middle of the screen and its angle is set
to 0. After that the light position is set to $\mathit{start}_i$ and the
position of the $\nkb $kilobots is set to $\mathit{start}_i + \mathit{off}_j$,
where $\mathit{off}_j$ is a small individual offset for kilobot number $j$.
Then $\nst$ time steps are simulated where in each one an action $a_k$ is chosen
for the current state $s_k$ according to $\pi$ and applied by adding $a_k$ to
the current light posiiton. Next the kilobots move directly toward the light and
lastly we record the current state as $s_k'$, compute the reward $r_k$ and
save the tuple $(s_k, a_k, r_k, s_k')$ as a sample. We also restrict the
magnitude of the light and kilobot movements to get a more realistic simulation.

The simulator can also be used to simulate the kilobots pushing an object
through a maze. In this mode we do not record any samples. Instead we move the
light according to our object movement policy which gets a target position from
our maze solving policy. The policies are combined in the following way:

As we will discuss in \autoref{sec:objpolicy} our object movement policy learns
how to move the light to push the object to the right given the current light
and kilobot positions relative to the object's position. The maze solving policy
now gives us a target position $t$ from which we compute the target direction as
$d = t - o$ where $o$ is the current object position. We mow compute the angle
$\alpha$ between $d$ and $(1, 0)$ (the vector pointing to the right) and rotate
the state around $\alpha$. Then we use the object movement policy to get an
action for our rotated state which we rotate by $-\alpha$ to get the final
action which we apply just as in the sample generation case.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.9\linewidth]{figures/simulator_maze.png}
    \caption{Simulator showing 4 kilobots pushing an object through a maze}
    \label{fig:simulator}
\end{figure}

\subsection{Object Movement Policy}
\label{sec:objpolicy}

% sample generation iteration
% states
% features

\section{RESULTS}

\section{FUTURE WORK \& CONCLUSION}
